---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# IMDB ratings dataset exploratory analysis

#4 based on : https://www.kaggle.com/vipluv07/imdb-visualization-with-r/notebook?select=Movie+Ratings.csv

I want to understand data science deeper and this is part of my learning journey.


```{r include=FALSE}
#Some preliminary things:
#install.packages(c("tidyverse", "ggplot2", "treemap"))
library(ggplot2)
library(treemap)
library(tidyverse)
library(cowplot)
backgroundColor = "#082021"
lineColor = "#617073"
colorPalette = c("#6b2737", "#e08e45", "#e0e2db", "#889696", "#1b2f33")
gradientPalette = c("#6b2737", "#783c49", "#85515b", "#92656e", "#9f7a80", "#ac8f92", "#b9a4a4", "#c6b8b7", "#d3cdc9", "#e0e2db")
```





```{r echo=FALSE}

#Let's import the data:
IMDB = read.csv("Movie-Ratings.csv")

```

First let's see what was imported:

```{r echo=FALSE}
head(IMDB)
summary(IMDB)
tail(IMDB)
str(IMDB)
```

With column names set nicely it is easier to work with the data:

```{r echo=FALSE}
colnames(IMDB) <- c("Title", "Genre", "CriticRating", "ViewerRating", "Budget", "Year")
head(IMDB)
```

Might make sense to check these (are genre and year factors in reality?):

```{r echo=FALSE}
table(IMDB$Genre)
table(IMDB$Year)
```

Checking the structure we can see that the Year and the Genre is actually categorical but it is threated as a number and a string by R, let's fix this:

```{r echo=FALSE}
IMDB$Year <-factor(IMDB$Year)
IMDB$Genre <- factor(IMDB$Genre)
str(IMDB)
```

With the data being nice and tidy it is time to analyze and visualize:

We have in our hands a sample of movies. The sample size is:

```{r echo=FALSE}
length(IMDB$Title)
```

We have movies from multiple years:

```{r echo=FALSE}
#plot(IMDB$Year)
ggplot(IMDB, aes(x=Year)) + 
        geom_bar(stat = "count", width = 0.6, fill = colorPalette[1]) + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))


```
And across many genres: 
```{r echo=FALSE, warning=FALSE}
pie(table(IMDB$Genre), col = gradientPalette, 
    main = "Genre Pie Chart", border = "white")

genres = as.character(unique(IMDB$Genre))
values = as.numeric(table(IMDB$Genre))
df = data.frame(genres, values)
treemap(df, index = "genres", vSize = "values",border.col=backgroundColor, border.lwds=1, title="Genre distributtion", 
        palette = gradientPalette, 
        fontsize.labels=0.5,
        fontcolor.labels="white",
        #fontface.labels=1,            
        bg.labels=c("transparent"),              
        align.labels=c("left", "top"),                                  
        overlap.labels=0.5,
        inflate.labels=T )


```

Would you trust the critics or the audience more? 

```{r}
ggplot(data = IMDB, aes(x=CriticRating, y = ViewerRating, colour = Genre, size = Budget)) + 
        geom_point() + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))

```

Is the money turned into high ratings? 

```{r}
rating1 <- ggplot(data = IMDB, aes(x=ViewerRating, y = Budget, colour = Genre)) + 
        geom_point() + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor), legend.position = "none")
rating2 <- ggplot(data = IMDB, aes(x=CriticRating, y = Budget, colour = Genre)) + 
        geom_point() + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor), legend.position = "none", axis.text.y = element_blank(), axis.title.y = element_blank())

plot_grid(rating1, rating2)
```

```{r}
ggplot(data = IMDB, aes(x = Budget)) + 
        geom_histogram(binwidth = 10, aes(fill = Genre), colour = "White") + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))

```

```{r}
plot1 <- ggplot(data = IMDB, aes(x = CriticRating)) + 
        geom_histogram(binwidth = 10, aes(fill = Genre), colour = "White") + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor), 
              legend.position = "none", axis.text.y = element_blank(), axis.title.y = element_blank())
plot2 <- ggplot(data = IMDB, aes(x = ViewerRating)) + 
        geom_histogram(binwidth = 10, aes(fill = Genre), colour = "White") + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor), 
              legend.position = "none", axis.text.y = element_blank(), axis.title.y = element_blank())
plot_grid(plot1, plot2)
```

```{r}
ggplot(data = IMDB, aes(x=CriticRating, y= ViewerRating, colour = Genre )) + geom_smooth(fill = NA) + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))

```

```{r}
plot1 <- ggplot(data = IMDB, aes(x=Genre, y= ViewerRating, colour = Genre )) + 
        geom_jitter() + geom_boxplot(alpha = 0.5) + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))
plot2 <- plot1 <- ggplot(data = IMDB, aes(x=Genre, y= CriticRating, colour = Genre )) + 
        geom_jitter() + geom_boxplot(alpha = 0.5) + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))
plot_grid(plot1, plot2, ncol = 1)
```

```{r}
ggplot(data=IMDB,aes(x=Budget)) + geom_histogram(binwidth=10,color='white',aes(fill = Genre)) + 
        facet_grid(Genre~.,scales = 'free') + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))
```

```{r}
ggplot(data = IMDB, aes(x = ViewerRating, y = CriticRating, colour = Genre)) +
        geom_point() + facet_grid(Genre~.) + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))

```

```{r message=FALSE, warning=FALSE}
ggplot(data = IMDB, aes(x= CriticRating , y = ViewerRating , colour = Genre)) +
        geom_point() + geom_smooth()  + facet_grid(Genre~Year) + coord_cartesian(ylim=c(0,100)) + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))

```

```{r}
ggplot(data = IMDB, aes(x = Budget)) + 
        xlab("Budget") + 
        ylab("Number of movies") + 
        ggtitle("Movie and Budget Relation") + 
        geom_histogram(binwidth =10, aes(fill = Genre), colour = "white") +
        theme(legend.position = c(1,1), legend.justification= c(1,1), plot.title= element_text(colour = "Blue", size = 30, family = "courier" )) + 
        theme(panel.background = element_rect(fill=backgroundColor, colour = lineColor))

```

